Section Rings of $\bq$-divisors on minimal rational surfaces

### 2017-11-14 surfaces-write-up/surface-paper.tex
surfaces: minor intro edits

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{section ring} is $R(X, D) := \bigoplus_{d \geq 0}
H^0(X, \lfloor dD \rfloor)$.
%, where $u$ is a dummy variable to keep track of degree.
In the case that $D = K_X$, the canonical divisor, the graded section ring is referred to as
``the canonical ring'' and is a classical object of study.
For example, if $C$ is a curve of genus $g \geq 4$, Petri's theorem
relates the geometry of the curve $C$ to the canonical ring:
$R(C, K_C)$ is generated in degree 1 with relations in degree 2 
unless $C$ is hyperelliptic, trigonal, or a plane quintic
(see
\cite[p. 157]{saint-donat:proj} and \cite[Section 3.3]{acgh:algebraic-curves}).
In this way, explicit descriptions of generators and relations of section rings yield
geometric information about the underlying variety.

One natural way to generalize of the classical result of Petri mentioned above
is to examine the section rings of stacky curves (i.e., smooth proper
geometrically connected 1-dimensional Deligne-Mumford
stacks over a field with a dense open subscheme).
Indeed,
these were studied by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring},
which provide tight bounds on the degree of generators and relations
of log canonical rings and log spin canonical rings on arbitrary
stacky curves.
All rings of modular forms associated to Fuchsian groups can be realized as canonical rings of
such curves, and so the above work also yields insight into such rings
of modular forms.
Further, 
O'Dorney \cite{dorney:canonical} gives similar descriptions
of section rings for arbitrary $\bq$-divisors on $\bp^1$
(as opposed to just log canonical $\bq$-divisors and log spin canonical $\bq$-divisors).

In addition to section rings of curves, section rings of certain higher dimensional
stacks have also been studied.
For example, the Hassett-Keel program \cite{hassett:classical-and-minimal-models}
studies log canonical rings on ${\mathscr M}_g$ of the form 
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0} H^0 \left(
	\overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} +
	\alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces, where $\overline {\mathscr M_g}$ is the
moduli space of stable genus $g$ curves.

@@

In this paper, we continue the study of section rings of
surfaces and higher dimensional varieties, examining section rings of
projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a presentation of the
section ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

>>>>

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{section ring} is $R(X, D) := \bigoplus_{d \geq 0}
H^0(X, \lfloor dD \rfloor)$.
%, where $u$ is a dummy variable to keep track of degree.
In the case that $D = K_X$, where $K_X$ is the canonical divisor,
the graded section ring is referred to as
``the canonical ring'' and is a classical object of study.
For example, if $C$ is a curve of genus $g \geq 4$, Petri's theorem
relates the geometry of the curve $C$ to the canonical ring:
$R(C, K_C)$ is generated in degree 1 with relations in degree 2 
unless $C$ is hyperelliptic, trigonal, or a plane quintic
(see
\cite[p. 157]{saint-donat:proj} and \cite[Section 3.3]{acgh:algebraic-curves}).
In this way, explicit descriptions of generators and relations of section rings yield
geometric information about the underlying variety.

One natural way to generalize of the classical result of Petri
mentioned above is to examine the section rings of stacky curves
(i.e., smooth proper geometrically connected 1-dimensional
Deligne-Mumford stacks over a field with a dense open subscheme).
These were studied by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring},
which provide tight bounds on the degree of generators and relations
of log canonical rings and log spin canonical rings on arbitrary
stacky curves. All rings of modular forms associated to Fuchsian
groups can be realized as canonical rings of
such curves, so the above work also yields insight into such rings
of modular forms.
Further, 
O'Dorney \cite{dorney:canonical} gives similar descriptions
of section rings for arbitrary $\bq$-divisors on $\bp^1$
(as opposed to just log canonical $\bq$-divisors and log spin canonical $\bq$-divisors).

Beyond section rings of curves, section rings of certain higher dimensional
stacks have also been studied.
For example, the Hassett-Keel program \cite{hassett:classical-and-minimal-models}
studies log canonical rings on ${\mathscr M}_g$ of the form 
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0} H^0 \left(
	\overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} +
	\alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces, where $\overline {\mathscr M_g}$ is the
moduli space of stable genus $g$ curves.

@@

In this paper, we continue the study of section rings of
surfaces and higher dimensional varieties, examining section rings of
projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on $\bp^m$ or
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a presentation of the
section ring when $D$ is any effective
$\bq$-divisor on $\bp^m$.
